Poker Chips, Earthquakes, and Computation

### 2014-10-11 finaldraft.tex
starting computation section

@@

\subsection{Model of Computation}

>>>>

\begin{figure}
\centering
\subfigure[Board before earthquake.]{\includegraphics[width=0.25\linewidth]{action1.pdf}\label{fig:boardbeforeearthquake}}
\qquad\qquad
\subfigure[Board after earthquake to active square]{\includegraphics[width=0.25\linewidth]{action2.pdf}\label{fig:boardafterequarthquake}}
\caption{Earthquake applied to board. Green signifies active nodes}
\label{fig:earthquakeexample}
\end{figure}

@@

\subsection{Simple Model}

A simple model of computation can be defined as follows. We assume a finite board. We specify some input nodes and some output nodes. If an input bit is to be set to 1, then we place 4 chips in the input node. We apply earthquakes to the board until it stabilizes. If an output node ever had an earthquake, then we mark that output bit as 1; otherwise, the output bit is set to 0.

A simple example is a board that computes the identity function of one output. We can compute 
\[ f(x) = x \]
By having a board with no chips, and we let the input node be the same as the output node. Then if $x = 0$, this means that the input node will not contain any chips, so the board is stable. The output node never received an earthquake, so the output is 0. If $x = 1$, then the input node gets 4 chips and an earthquake affects it. So the output node is activated, and so the output is 1.
With this simple model, there are a surprising number of things we can do. We can make wires, AND gates and OR gates (see Figure~\ref{fig:simplegates}).

\begin{figure}
\centering
\subfigure[Wire]{\includegraphics[width=0.25\linewidth]{wire.pdf}\label{fig:wire}}
\qquad\qquad
\subfigure[AND gate]{\includegraphics[width=0.25\linewidth]{andgate.pdf}\label{fig:AND}}
\qquad\qquad
\subfigure[OR gate]{\includegraphics[width=0.25\linewidth]{orgate.pdf}\label{fig:OR}}
\caption{Gates in the simple model. Green are the inputs, red are the outputs}
\label{fig:simplegates}
\end{figure}

Composing these gates, we can make circuits which represent some Boolean function. In particular we can compute any Boolean function that can be represented with a planar circuit with only AND and OR.

\begin{theorem}
There does not exists a board which computes $f(x) = \neg x$.
\end{theorem}

\begin{proof}
Suppose for contradiction that we have a board $b$ with one input $i$ and one output $o$ which computes NOT. Then we know that there exists some node which initially has over $4$ chips, since otherwise, the board would be stable and so the output never gets activated. 
\end{proof}
